---
title: "Dataprep"
output: html_document
date: "2024-09-25"
---

```{r libraries, include = F}
#rm(list = ls())
require(tidyverse)
require(igraph)
require(shiny)
require(RSiena)
require(rmdformats)
require(prettydoc)
require(hrbrthemes)
require(tint)
require(tufte)
require(psych)
```

```{r setup, include=FALSE}
load("/Users/hannah/Desktop/Season 7, Part 1/Social Networks/labjournal/scholars_20240924.rda")
scholars <- x
rm(x)
```

# Chunk 1

```{r copied from jochem 1}
#copied from jochem
# (sociology, RU)
demographics <- do.call(rbind.data.frame, scholars$demographics)
demographics <- demographics %>% # who's who
    mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
        is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
        ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
        is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

sample <- which((demographics$Universiteit1.22 == "RU" | demographics$Universiteit2.22 == "RU" | demographics$Universiteit1.24 ==
    "RU" | demographics$Universiteit2.24 == "RU") & (demographics$discipline.22 == "sociology" | demographics$discipline.24 ==
    "sociology")) #sociologists at ru

demographics_soc <- demographics[sample, ]
scholars_sel <- lapply(scholars, "[", sample)
```

# Chunk 2

```{r copied from jochem 2}
ids <- demographics_soc$au_id #openalex ids
wave2 <- wave1 <- matrix(0, nrow = length(ids), ncol = length(ids), dimnames = list(ids, ids))
#matrix for all ids
```

# Chunk 3

```{r copied from jochem 3}
works <- scholars_sel$work
works_id <- unlist(lapply(works, function(l) l$id)) #id for works
works_author <- unlist(lapply(works, function(l) l$author), recursive = FALSE) #authors of the work
works_year <- unlist(lapply(works, function(l) l$publication_year), recursive = FALSE) #year
df_works <- tibble(works_id, works_author, works_year) #id-work, authors, year
view(df_works)
dups <- which(duplicated(works_id))
# why are there some many papers of Batenburg duplicates

df_works <- df_works[-dups, ]
df_works_newer <- df_works[df_works$works_year > 2014, ]
df_works_w2 <- df_works[df_works$works_year > 2019, ] 
```
```{r jochem deed dit, include = F, echo = F}
# Chunk 4

## If we wanted to, we could

#first to others
for (i in 1:nrow(df_works_w2)) {
    ego <- df_works_w2$works_author[i][[1]]$au_id[1]
    alters <- df_works_w2$works_author[i][[1]]$au_id[-1]

    if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
        wave2[which(ids %in% ego), which(ids %in% alters)] <- 1
    }
} 
#for each row, of dataframes of work, do egos (first auth) and alters (others) on matrix (and then: are they the same or not)

#last to others
for (i in 1:nrow(df_works_w2)) {
    ego <- rev(df_works_w2$works_author[i][[1]]$au_id)[1]
    alters <- rev(df_works_w2$works_author[i][[1]]$au_id)[-1]

    if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
        wave2[which(ids %in% ego), which(ids %in% alters)] <- 1
    }
} 
#for each row, of dataframes of work, do egos (last auth) and alters (others) on matrix (and then: are they the same or not)

#undirected
for (i in 1:nrow(df_works_w2)) {
    egos <- df_works_w2$works_author[i][[1]]$au_id

    if (sum(ids %in% egos) > 0) {
        wave2[which(ids %in% egos), which(ids %in% egos)] <- 1
    }
}
#for each row, of dataframes of work, do egos (all authors) on matrix (and then: are they the same or not)
```


# About the parameters:

[data]: our scholars file\
university: Character vector with names of universities. We have several universities in the Netherlands. See above for relevant names.\
discipline: Character vector, either sociology or political science or both.\
waves: a list of numeric vectors with start and end year of wave.\
type:\
- "first": directed: first author sending to others\
- "last": directed: last author sending to others\
- "all": undirected: ties between all authors\
Output: - a list\
- $nets: array of nomination networks.\
- $data: sample of data (scholars)\

# Chunk 5
```{r copied from jochem 5}
#so this is all we are doin'
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}

```
# Chunk 6

```{r}
#save the output of your function
matrix_ru_soc <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = "sociology", 
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))

mrs_graph1 <- igraph::graph_from_adjacency_matrix(
  matrix_ru_soc$nets[1,,], #for this example I take the first wave of data. (thus I select the array of networks and take the first matrix)
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL
)

plot(mrs_graph1,
  vertex.label= NA,
  edge.width = 0.2,
  edge.arrow.size =0.2)
```

# Chunk 7

more!
```{r}
matrix_ru_socpol <- fcolnet(data = scholars, 
                university = c("RU", "UvT"), 
                discipline = c("sociology", "political science"), 
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("all"))

mrsp <- igraph::graph_from_adjacency_matrix(
  matrix_ru_socpol$nets[2,,], #now, I take the second wave
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL
)

plot(mrsp,
     vertex.label = NA,
     vertex.size = .5,
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)
```

```{r now some more}
#Let us find ego characteristics. 
#first fish out the data
matrix_ru_socpol_df <- matrix_ru_socpol$data

#same complicated structure as 'scholars' thus first make a dataframe from the list in which all info was saved. 
mrusp_df_ego <- do.call(rbind.data.frame, matrix_ru_socpol_df$demographics)

#DO NOT MESS UP THE ORDER! THUS IF YOU JOIN THIS DATA WITH YOUR OWN DATA CHECK THAT ORDER REMAINED THE SAME!! 

plot(mrsp,
  vertex.color = ifelse(mrusp_df_ego$discipline.24 == "sociology", "red", "blue"), #now, I can use actor attributes for plotting. 
  vertex.label = NA,
  edge.width = 0.2,
  edge.arrow.size =0.2,
  vertex.size = 5)
```


# Hannah's Shit 

Ok, we want an overview of the different universities we are dealing with, and the different disciplines -- in other words, we want to see how clean our data actually is.\

```{r, include = F, echo = F}
universities_1_2022 <- unique(demographics$Universiteit1.22)
universities_2_2022 <- unique(demographics$Universiteit2.22)
universities_1_2024 <- unique(demographics$Universiteit1.24)
universities_2_2024 <- unique(demographics$Universiteit2.24)

universities_1_2022 
universities_2_2022
universities_1_2024
universities_2_2024

#"UU", "RUG", "UvA", "EUR", "Leiden", "Bocconi University", RU", "Universita degli studi di Milano", "VU", "UvT", "Universiteit Gent", "Universiteit Stockholm", "University Linköping", "TU Delft", "University of Turku", "Frankfurt School of Finance & Management", "University of Lausanne", "UvH", "Boston University", "Trento University", "Uni Gothenburg", "Uni Berlijn", "Uni of Victoria", "Uni of Lausanne", "Shadona uni", "University of Cologne","Shandong uni"   
```
The Universities\


|Abbreviation | Name and City | Country |
|:---------:|:---------:|:---------:|
|"UU"| Utrecht University | the Netherlands |
|"RUG"| Rijksuniversiteit Groningen | the Netherlands|
|"UvA"| Universiteit van Amsterdam | the Netherlands|
|"EUR"| Erasmus Universiteit Rotterdam | the Netherlands |
|"Leiden"| Leiden University | the Netherlands |
|"Bocconi University"| Bocconi University, Milan | Italy |
| RU" | Radboud University | the Netherlands |
|"Universita degli studi di Milano"| Universita degli studi di Milano | Italy |
|"VU"| Vrije Universiteit, Amsterdam | the Netherlands |
|"UvT"| Tilburg University | the Netherlands |
|"Universiteit Gent"| University Gent| Belgium |
|"Universiteit Stockholm"| University of Stockholm | Sweden |
|"University Linköping"| University of Linköping | Sweden |
|"TU Delft"| TU Delft | the Netherlands |
|"University of Turku"| University of Turku | Finland |
|"Frankfurt School of Finance & Management"| Frankfurt School of Finance & Management | Germany |
|"University of Lausanne"| University of Lausanne | Switzerland |
|"UvH"| University of Humanistic Studies, Utrecht | the Netherlands |
|"Boston University"| Boston University | the US |
|"Trento University"| Trento Univeristy | Italy |
|"Uni Gothenburg"| University of Gothenburg | Sweden |
|"Uni Berlijn"| University of Berling | Germany |
|"Uni of Victoria"| University of Victoria | Canada |
|"Uni of Lausanne"| University of Lausanne | Switzerland |
|"Shadona uni"| Shadong University | China | 
|"University of Cologne"| University of Cologne | Germany |
|"Shandong uni"| Shadong University | China |
\

Now see if we can put this into a social network plot/graph/whatever. 

```{r}
matrix_all_socpol <- fcolnet(data = scholars, 
                university = c("UU", "RUG", "UvA", "EUR", "Leiden", "RU", 
                               "VU", "UvT", "TU Delft", "UvH"), 
                discipline = c("sociology", "political science"), 
                waves = list(c(2015, 2018), c(2019, 2023))) 

mall_w1 <- igraph::graph_from_adjacency_matrix(
  matrix_all_socpol$nets[1,,],
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL
)

#nicely_chosen <- layout_nicely(mall_w1)
#kk <- layout_with_kk(mall_w1)
#lgl <- layout_with_lgl(mall_w1)
random <- layout_randomly(mall_w1)

plot(mall_w1,
     vertex.color = ifelse(mrusp_df_ego$discipline.24 == "sociology", "red", "blue"),
     vertex.size = 4,  # we'll vertices a bit smaller
     vertex.label = NA,  
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)
```
\
OK, so that worked! -- I think that if I add my gender variables (spoilers! for the next part) to this, I'll also be able to use that to create some fun. 

And extract all names of the researchers

```{r}
view(demographics)
names <- demographics$Naam
au_id <- demographics$au_id
names <- cbind(names, au_id)
view(names)
```












