---
title: "We've Got Graphs at Home"
output: html_document
date: "2024-09-29"
---
---
```{r setup, include=FALSE, echo = F}
knitr::opts_chunk$set(echo = TRUE)
```
```{r libraries, include=FALSE, echo = F}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(igraph)
require(shiny)
require(RSiena)
require(rmdformats)
require(prettydoc)
require(hrbrthemes)
require(tint)
require(tufte)
require(psych)
```
```{r first matrix, include=FALSE, echo = F}
matrix <- matrix(c(0,0,0,1,
                   1,0,0,0,
                   0,0,1,0,
                   0,0,0,0), nrow = 4, ncol = 4)
graph <- graph_from_adjacency_matrix(matrix)
matrix_graph <- triad_census(graph)
matrix_graph
```
```{r all matrix, include=FALSE, echo = F}
t1 <- s501 #matrix at t1
t2 <- s502 #matrix at t2
dim(s501) #dimensions full matrix
small_t1 <- s501[c(1:10), c(1:10)] #sample matrix t1
small_t2 <- s502[c(1:10), c(1:10)] #sample matrix t2
small_graph_t1 <- graph_from_adjacency_matrix(small_t1)
small_graph_t2 <- graph_from_adjacency_matrix(small_t2)
plot(small_graph_t1) #graph of sample matrix t1
plot(small_graph_t2) #graph of sample matrix t2, some broken relations I see

```
```{r small density, include=FALSE, echo = F}
density_t1 <- edge_density(small_graph_t1)
density_t1 #so that's approximately 11%
```
```{r small degree centrality, include=FALSE, echo = F}
degree_t1 <- degree(small_graph_t1, mode = "out") #only look at outgoing edges 
degree_t1 #not normalised, it seems 
degree_t1_norm_a <- degree(small_graph_t1, mode = "out", normalized = T)
degree_t1_norm_a #normalised using only sample? 

#let's see how it differs if we calculate degree according to the formula given by Jochem
normalised_cent_t1 <- (degree_t1 - min(degree_t1))/(max(degree_t1)) - min(degree_t1)
normalised_cent_t1 #significanly different, it appears

```
```{r small closeness centrality, include=FALSE, echo = F}
closeness_t1 <- closeness(small_graph_t1, mode = "out")
closeness_t1
```
```{r small betweenness centrality, include=FALSE, echo = F}
between_t1 <- betweenness(small_graph_t1, directed = T, normalized = F )
between_t1 #zeroes
between_t1_norm <- betweenness(small_graph_t1, directed = T, normalized = T)
between_t1_norm #normalisation does not matter, because the zeroes 
```
```{r small clustering, include=FALSE, echo = F}
transitivity_t1 <- transitivity(small_graph_t1, type = "local")
transitivity_t1
```
```{r large graphs, include=FALSE, echo = F}
graph_t1 <- graph_from_adjacency_matrix(t1)
plot.igraph(graph_t1, main = "Network at t1")
graph_t2 <- graph_from_adjacency_matrix(t2)
plot.igraph(graph_t2, main = "Network at t2")
```
```{r all density t1, include=FALSE, echo = F}
density_all_t1 <- edge_density(graph_t1)
density_all_t1
```
```{r all density t2, include=FALSE, echo = F}
density_all_t2 <- edge_density(graph_t2)
density_all_t2
```
```{r all degree centrality t1, include=FALSE, echo = F}
degree_all_t1 <- degree(graph_t1, mode = "out")
degree_all_t2 <- degree(graph_t2, mode = "out")

normalised_all_cent_t1 <- (degree_all_t1 - min(degree_all_t1))/(max(degree_all_t1)) - min(degree_all_t1)
hist(normalised_all_cent_t1, main = "Histogram Network Centrality t1", xlab = "Normalised Centrality")
```
```{r all degree centrality t2, include=FALSE, echo = F}
normalised_all_cent_t2 <- (degree_all_t2 - min(degree_all_t2))/(max(degree_all_t2)) - min(degree_all_t2)
hist(normalised_all_cent_t2, main = "Histogram Network Centrality t2", xlab = "Normalised Centrality")
```
```{r all closeness centrality t1 plot, include=FALSE, echo = F}
closeness_all_t1 <- closeness(graph_t1, mode = "out")
plot(closeness_all_t1, main = "Plot Closeness Centrality t1", ylab = "Normalised Closeness Centrality")
```
```{r all closeness centrality t1 histo, echo=F, include=FALSE, echo = F}
hist(closeness_all_t1, breaks = 100, main = "Histogram Closeness Centrality t1", xlab = "Normalised Closeness Centrality")
```
```{r all closeness centrality plot t2, include=FALSE, echo = F}
closeness_all_t2 <- closeness(graph_t2, mode = "out")
plot(closeness_all_t2, main = "Plot Closeness Centrality t2", ylab = "Normalised Closeness Centrality")
```
```{r all closeness centrality hist t2, include=FALSE, echo = F}
hist(closeness_all_t2, breaks = 100, main = "Histogram Closeness Centrality t2", xlab = "Normalised Closeness Centrality")
```
```{r all betweenness centrality t1, include=FALSE, echo = F}
between_all_t1 <- betweenness(graph_t1, directed = T)
hist(between_all_t1, breaks = 100, main = "Histogram Betweenness Centrality t1", xlab = "Betweenness Centrality")
```
```{r all betweenness centrality t2, include=FALSE, echo = F}
between_all_t2 <- betweenness(graph_t2, directed = T)
hist(between_all_t2, breaks = 100, main = "Histogram Betweenness Centrality t2", xlab = "Betweenness Centrality")
```
```{r all clustering global, include=FALSE, echo = F}
transitivity_all_t1 <- transitivity(graph_t1, type = "global")
transitivity_all_t1
transitivity_all_t2 <- transitivity(graph_t2, type = "global")
transitivity_all_t2
```
```{r all clustering local t1, echo=F, include=FALSE, echo = F}
transitivity_t1 <- transitivity(graph_t1, type = "local")
hist(transitivity_t1, breaks = 20, main = "Histogram Clustering t1", xlab = "Clustering t1")
```
```{r all clustering local t2, include=FALSE, echo = F}
transitivity_t2 <- transitivity(graph_t2, type = "local")
hist(transitivity_t2, breaks = 20, main = "Histogram Clustering t2", xlab = "Clustering t2")
```

# How To Plot a Network

Until now, we have worked with the standard settings of the igraph plot feature. Now we're done with the boring stuff (descriptives (which are of course very interesting of course)), we can take a look at the different ways in which we can adjust graphs, and make them look different and more to our liking. 

## igraph
Some surfing on the web brings us [this](https://kateto.net/network-visualization) gem of a website from Katya Ognyanova. It contains an impressively extensive number of ways in which one can create and modify networks in R. 
To be able to follow the instructions from that website, we first `require` the relevant packages 
```{r libraries kateto, echo=T, message=F}
require(igraph) 
require(network)
require(sna)
require(ggraph)
require(visNetwork)
require(threejs)
require(networkD3)
require(ndtv)
```
Then we can extract some attributes from our graph objects to increase the ease of use of packages other than `igraph` to draw and plot networks. 
```{r extract graph info}
edges_t1 <- E(graph_t1) #extract edges from t1    
vertices_t1 <- V(graph_t1) #extract vertices from t1 
```

The first plot we'll draw is a similar one to those which we have plotted before: a simple `igraph` network plot with no added constraints. 
```{r simpeles plot igraph}
plot.igraph(graph_t1)
```

This is not exactly the most legible graph, but we can change that! We do this by introducing several constraints that eliminate some of the clutter: we remove loops and vertice labels, and decrease the size of the arrows.
```{r remove labels}
plot.igraph(graph_t1, remove.loops = T, edge.arrow.size = .4, vertex.label = NA)
```

Then, since the edges themselves are still not as easily distinguishable as we would like them to be, we modify the edges and make them somewhat curved.
```{r do some curved edges}
plot.igraph(graph_t1, remove.loops = T, edge.arrow.size = .4, vertex.label = NA, edge.curved=.5)
```

That helped ... somewhat. Lets see if there are any further characteristics included in the matrices we have been using. 
```{r see if there are any more kinds of info here}
str(s501)
```

OK so there are absolutely no other kinds of information here --- which is sad for those of us who'd like more information to be included in the plot (like me), but despair not, there is more to do with what we have!
We can, for example, adjust the layout through some functions included in the package. 
```{r several layouts testing}
graph_attr(graph_t1, "layout") <- layout_on_grid #layout on a grid
plot(graph_t1) 
graph_attr(graph_t1, "layout") <- layout_with_graphopt #layout with force-directed algoritm
plot(graph_t1) #I think that we need a way to have the graph be less clumped, so we can better understand what we see
graph_attr(graph_t1, "layout") <- layout_with_fr #force directed
plot(graph_t1) 
```

The main problem seems to be that the vertices are way too big to see what is 'under there'. So let us see if we can solve *that* issue next. 
```{r random layout using these things}
random <- layout_randomly(graph_t1)
plot(graph_t1, layout=random) #first we make a random layout
#so what does size 5 give us?
plot.igraph(graph_t1, layout=random, remove.loops = T, edge.arrow.size = .4, vertex.label = NA, edge.curved=.5, vertex.size = 5) #a way better result -- maybe see if even smaller would help, but probably not
plot.igraph(graph_t1, layout=random, remove.loops = T, edge.arrow.size = .4, vertex.label = NA, edge.curved=.5, vertex.size = 1) #the very fancy thing that's happening here, is that we've moved on from focussing on vertices, to focussing on edges. 'Cause that's all that's showing up here. 
```

So now we have a way to make the connections --- the edges --- more legible, though also at the cost of the bottom edge, but that's something I can currently live with. The next step is highlighting the communities in the graph. The following chunck is based on code from [Ognyanova](https://kateto.net/network-visualization)
```{r copied from ognyanova look at her}
# First we 'detect' the optimal clustering, i.e. the groups in the sample
clp <- cluster_optimal(graph_t1)
class(clp)

# then, when we have detected the communities, we include them in the plot syntax, to yield a plot based on these communities
plot(clp, graph_t1, remove.loops = T, edge.arrow.size = .4, vertex.label = NA, edge.curved=.5, vertex.size = 1) #baby that looks fun!
```

We can also do it ourselves, by assigning colours to the vertices based on the communities found in the graph.
```{r do it ourselves}
V(graph_t1)$community <- clp$membership
#creating our own set of colours
colrs <- adjustcolor( c("gray50", "tomato", "gold", "yellowgreen"), alpha=.6)
#and making a graph
plot(graph_t1, remove.loops = T, edge.arrow.size = .4, vertex.label = NA, edge.curved=.5, vertex.size = 10, vertex.color=colrs[V(graph_t1)$community]) #had to make the vertices somewhat larger to make the colour legible, but this also looks fun --- though the circle-things do have my preference
```

Due to the limitations of our data --- it only yielding information on ties, and nothing on characteristics of the nodes --- there's not much else we can do. However, there are other packages we can use to do some more exiting things `igraph` cannot --- as far as I know --- do (such as interactive graphs!)

\

## ggraph
Love of my life ggplot also --- of course --- has a way to graph networks. To see how much we can adjust the plot to our likings we are gonna fuck around and find out, by first doing something very simple (because to find out by fucking around, we must first know what we are fucking around with).
```{r ggraph simple}
ggraph(graph_t1) + #the basic for the graph-object 
  geom_edge_link() +  #edges
  geom_node_point()   #nodes 
```

\
Boy is she ugly, let's see if we can solve that:
```{r}
ggraph(graph_t1, layout = "fr") + #force directed?
  geom_node_point() +
  geom_edge_arc(color="gray", strength = 0.3) + 
  theme_void()
#Rorschachtest for network analysts: what do you see? (hint: happy feet?)
```
\
Ggraph can do much more, especially if your edges and vertices have (multiple) characteristics. One of the main advantages of ggplot/graph in this respect is that the legend is automatically included, and that you can quite easily adjust the size of the graphs, which is something I have struggled with in `igraph`. 

\

## visNetwork

`visNetwork` gives you graphs that you can zoom in- and out- of, and can 'look around' in. While this is more fun when your graph is much bigger, and more informative, it might be a better idea to learn how it works on a smaller network you already have some experience with. As such, let us commence. 

The plotting functions in `visNetwork` work with seperate objects representing the vertices and edges of a graph, instead of the `igraph` objects we have thus far worked with. Luckily enough, there's a simple way to convert the latter into an object the former can work with. 
```{r interactive visNetwork setup part 1, include=F}
#first we look at the two objects, to make sure we can use them in the visNetwork function 
visnet_t1 <- toVisNetworkData(graph_t1)
summary(visnet_t1)
```
```{r interactive visNetwork output part 1}
visnet_t1_vertices <- visnet_t1$nodes
visnet_t1_edges <- visnet_t1$edges
visNetwork(visnet_t1_vertices, visnet_t1_edges, width="100%", height="400px") #look at her i can move things around!
```
\
With some adjustments to the nodes and links, we can make them a bit sexier --- add some dropshadows, some names, and if we had the info, we could have even added some edge-weight and vertix-size values related to scores and values. 

## Configurations 
`igraph` can also do some other things for the looks of our plot, ones that are even *functional*. One of these is the layout function, which includes several different algorithmic functions with which to configure the layout of our plots. 

```{r random layouts all igraph}
random <- layout_randomly(graph_t1) #because we're not interested in descriptives anymore, we're just dealing with the t1 graph. 
plot(graph_t1, layout = random) #a random layout

circle <- layout_in_circle(graph_t1) #a circle
plot(graph_t1, layout = circle) #well, that's an absolute choice i guess... 

sphere <- layout_on_sphere(graph_t1) #a sphere
plot(graph_t1, layout = sphere) #well that just tickles my fancy, absolutely useless for yielding any information, but so fun to look at

drl <- layout_with_drl(graph_t1) #Distributed Recursive Layout
plot(graph_t1, layout = drl) #this i Hate, absolutely Hate

fr <- layout_with_fr(graph_t1) #Fruchterman-Reingold force-directed algorithm
plot(graph_t1, layout = fr) #Honestly, looks quite ok, not as fun to look at as some others, but more informative.

lgl <- layout_with_lgl(graph_t1) #The LGL (Large Graph Layout) algorithm
plot(graph_t1, layout = lgl) #so only the big chunk of interconnected nodes shows up

kk <- layout_with_kk(graph_t1) #Kamada-Kawai force-directed algorithm
plot(graph_t1, layout = kk) #fun, informative, cool, hip, it's awesome (especially that half moon on the right, big fun)

nicely_chosen <- layout_nicely(graph_t1) #nice
plot(graph_t1, layout = nicely_chosen) #nice 
#also: mirrored Fruchterman-Reingold?
```

\

## Colours 

There's also some further bullshit we can employ to add some sparkle to our plots.

```{r colour nodes}
#we are, of course, starting with a nice graph
plot(graph_t1, layout = nicely_chosen, vertex.color = "#A6BCB0") #and now it's sage bb
plot(graph_t1, layout = nicely_chosen, vertex.color = "#A6BCB0", vertex.label.color = "black") #lovely, changed the text
plot(graph_t1, layout = nicely_chosen, vertex.color = "#A6BCB0", vertex.label.color = "black", edge.color = "black") #got me some black edges too
plot(graph_t1, layout = nicely_chosen, vertex.color = "#A6BCB0", vertex.label.color = "black", edge.color = "black", edge.curved = T) #ok so thus arrows are doing me wrong right now
plot(graph_t1, layout = nicely_chosen, vertex.color = "#A6BCB0", vertex.label.color = "black", edge.color = "black", edge.curved = T, edge.arrow.size = 0.3, size = 50, vertex.label = NA) #have solved the arrows, life is good again, now have to find a way to get those vertices further away from each other
plot(graph_t1, vertex.color = "#A6BCB0", vertex.label.color = "black", edge.color = "black", edge.curved = T, edge.arrow.size = 0.3, size = 50, vertex.label = NA) #maybe the real friends were the layout we had all along
```
