---
title: "RSiena Analysis"
output: html_document
date: "2024-10-16"
---
# Get the Basics
```{r packages, message = F, warning=F}
require(tidyverse)
require(igraph)
require(shiny)
require(RSiena)
require(rmdformats)
require(prettydoc)
require(hrbrthemes)
require(tint)
require(tufte)
require(psych)
library(data.table)  
library(xml2)
library(rvest)
library(xml2)
library(XML)
library(labelled)
library(forcats)
```

```{r dataset, include=F, warning=F, message=FALSE}
wave_1 <- read_rds("datasets_final/wave_1_first_authors.rds")
wave_2 <- read_rds("datasets_final/wave_2_first_authors.rds")
demographics_hannah <- read_csv("demographics_hannah_final.csv", show_col_types = FALSE)
view(demographics_hannah)
net_soc_array <- read_rds("datasets_final/net_soc_array_first_auth.rds")
```

```{r fcolnet, include = F, warning=F, message=F}
fcolnet <- function(data = scholars, university = "RU", 
                    discipline = "sociology", waves = list(c(2015,
                                                             2018), c(2019, 2023)), type = c("first")) {
  
  # step 1
  demographics <- do.call(rbind.data.frame, data$demographics)
  demographics <- demographics %>%
    mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
                                                                                                                 is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
                                                                                                                                                                          ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
                                                                                                                                                                                                                                                                                  is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))
  
  sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
                     university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
                     university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))
  
  demographics_soc <- demographics[sample, ]
  scholars_sel <- lapply(scholars, "[", sample)
  
  # step 2
  ids <- demographics_soc$au_id
  nwaves <- length(waves)
  nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
                                                                              ids))
  dimnames(nets)
  
  # step 3
  df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
                                                                                                                  function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
                                                                                                                                                                                 recursive = FALSE))
  
  df_works <- df_works[!duplicated(df_works), ]
  
  # step 4
  if (type == "first") {
    for (j in 1:nwaves) {
      df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
      ]
      for (i in 1:nrow(df_works_w)) {
        ego <- df_works_w$works_author[i][[1]]$au_id[1]
        alters <- df_works_w$works_author[i][[1]]$au_id[-1]
        if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
          nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
        }
      }
    }
  }
  
  if (type == "last") {
    for (j in 1:nwaves) {
      df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
      ]
      for (i in 1:nrow(df_works_w)) {
        ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
        alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
        if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
          nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
        }
      }
    }
  }
  
  if (type == "all") {
    for (j in 1:nwaves) {
      df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
      ]
      for (i in 1:nrow(df_works_w)) {
        egos <- df_works_w$works_author[i][[1]]$au_id
        if (sum(ids %in% egos) > 0) {
          nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
        }
      }
    }
  }
  output <- list()
  output$data <- scholars_sel
  output$nets <- nets
  return(output)
}

```

```{r user defined functions, include = F, message=F, warning=F}
GeodesicDistribution <- function(i, data, sims, period, groupName, varName, levls = c(1:5, Inf), cumulative = TRUE,
    ...) {
    x <- networkExtraction(i, data, sims, period, groupName, varName)
    require(sna)
    a <- sna::geodist(symmetrize(x))$gdist
    if (cumulative) {
        gdi <- sapply(levls, function(i) {
            sum(a <= i)
        })
    } else {
        gdi <- sapply(levls, function(i) {
            sum(a == i)
        })
    }
    names(gdi) <- as.character(levls)
    gdi
}

# The following function is taken from the help page for sienaTest

testall <- function(ans) {
    for (i in which(ans$test)) {
        sct <- score.Test(ans, i)
        cat(ans$requestedEffects$effectName[i], "\n")
        print(sct)
    }
    invisible(score.Test(ans))
}
```

# Variables 

```{r variables definition}
#dependent variable 
net <- sienaDependent(net_soc_array)

#independent variable 
hoogleraar_covar <-  coCovar(demographics_hannah$hoogleraar)
functie_covar <- coCovar(demographics_hannah$functie)
gender_covar <- coCovar(demographics_hannah$perc_female)
country_covar <- coCovar(demographics_hannah$born_in_binary)
white_pass_covar <- coCovar(demographics_hannah$white_or_not)
h_index_covar <- coCovar(demographics_hannah$h_index)
```

# Get Effect Structure 

```{r good data}
the_data <- sienaDataCreate(net, white_pass_covar, country_covar, gender_covar, functie_covar, h_index_covar)
myeff_1 <- getEffects(the_data)
myeff_1
```

```{r where to put it, include=F, echo=F}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)
print01Report(the_data, modelname = "./results/prelimsh*t")  
```
# Models - Parameter Effects {.tabset}
## M1: Race

### Making the Model
```{r}
myeff_1 <- getEffects(the_data)
view(myeff_1)
myeff_1 <- includeEffects(myeff_1, sameX, interaction1 = "white_pass_covar")
myeff_1 <- includeEffects(myeff_1, egoX, altX, interaction1 ="white_pass_covar")
myeff_1 <- includeEffects(myeff_1, inpopX, interaction1 ="white_pass_covar")
```

### Results
```{r results alogoritm, echo = F}
myAlgorithm <- sienaAlgorithmCreate(projname = "preliminary")
mod_1_effects <- siena07(myAlgorithm, data = the_data, effects = myeff_1, returnDeps = TRUE)
mod_1_effects 
```

```{r, include = F, echo = F}
#mod_1_effects <- siena07(myAlgorithm, data = the_data, effects = myeff_1, prevans = mod_1_effects, returnDeps=TRUE)
#?siena07
# ans_M2_2
```

## M2: Ethnicity

```{r model 2}
myeff_2 <- getEffects(the_data)
myeff_2 <- includeEffects(myeff_2, sameX, interaction1 = "country_covar")
myeff_2 <- includeEffects(myeff_2, egoX, altX, interaction1 = "country_covar")
myAlgorithm <- sienaAlgorithmCreate(projname = "preliminary")
mod_2_effects <- siena07(myAlgorithm, data = the_data, effects = myeff_2, returnDeps = TRUE)
mod_2_effects
#write_rds(mod_2_effects, "mod_2_effecs_conv_1.rds")
```


## M3: Gender

```{r gender model}
myeff_3 <- getEffects(the_data)
myeff_3 <- includeEffects(myeff_3, sameX, interaction1 = "gender_covar")
myeff_3 <- includeEffects(myeff_3, egoX, altX, interaction1 = "gender_covar")
myAlgorithm <- sienaAlgorithmCreate(projname = "preliminary")
mod_3_effects <- siena07(myAlgorithm, data = the_data, effects = myeff_3, returnDeps = TRUE)
mod_3_effects
#write_rds(mod_3_effects, "mod_3_effecs_conv_1.rds")
```



## M4: Gender and Race
```{r fourthmodel, include = T}
myeff_4 <- includeEffects(myeff_1, sameX, interaction1 = "gender_covar")
myeff_4 <- includeEffects(myeff_4, egoX, altX, interaction1 = "gender_covar")
myAlgorithm <- sienaAlgorithmCreate(projname = "preliminary")
mod_4_effects <- siena07(myAlgorithm, data = the_data, effects = myeff_4, returnDeps = TRUE)
mod_4_effects
#write_rds(mod_4_effects, "mod_4_effecs_conv_1.rds")
```
## M5: Gender and Ethnicity

```{r fifththmodel, include = T}
myeff_5 <- includeEffects(myeff_2, sameX, interaction1 = "gender_covar")
myeff_5 <- includeEffects(myeff_5, egoX, altX, interaction1 = "gender_covar")
myAlgorithm <- sienaAlgorithmCreate(projname = "preliminary")
mod_5_effects <- siena07(myAlgorithm, data = the_data, effects = myeff_5, returnDeps = TRUE)
mod_5_effects
#write_rds(mod_5_effects, "mod_5_effecs_conv_1.rds")
```

# GOF -- Parameter Models {.tabset}

## M1

### Create Effects
```{r GOF functions Model 1, message = F, warning=F}
model_1_effects <- read_rds("mod_1_effects_conv_1.rds")
gofi0_1 <- sienaGOF(model_1_effects, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "net")
gofo0_1 <- sienaGOF(model_1_effects, OutdegreeDistribution, verbose = FALSE, join = TRUE, levls = c(0:10, 15, 20),
    varName = "net")
gof0.gd_1 <- sienaGOF(model_1_effects, GeodesicDistribution, cumulative = FALSE, verbose = FALSE, join = TRUE, varName = "net")
```

### Describe Effects
```{r GOF Graphs Model 1}
plot(gofi0_1, main = "")
plot(gofo0_1, main = "")
plot(gof0.gd_1, main = "") #so these don't work, we need to adjust! (like, with other models? so look at 4)
```

## Model 4

### Create Effects
```{r GOF functions Model 4}
model_4_effects <- read_rds("mod_4_effecs_conv_1.rds")
gofi0_4 <- sienaGOF(model_4_effects, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "net")
gofo0_4 <- sienaGOF(model_4_effects, OutdegreeDistribution, verbose = FALSE, join = TRUE, levls = c(0:10, 15, 20),
    varName = "net")
gof0.gd_4 <- sienaGOF(model_4_effects, GeodesicDistribution, cumulative = FALSE, verbose = FALSE, join = TRUE, varName = "net")
```

### Describe Effects
```{r GOF Graphs Model 4}
plot(gofi0_4, main = "") # still a tragedy bb
plot(gofo0_4, main = "") #horrible
plot(gof0.gd_4, main = "") #bad (aka: need structural effects)
```

## Model 5

### Create Effects
```{r GOF functions Model 5}
model_5_effects <- read_rds("mod_5_effecs_conv_1.rds")
gofi0_5 <- sienaGOF(model_5_effects, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "net")
gofo0_5 <- sienaGOF(model_5_effects, OutdegreeDistribution, verbose = FALSE, join = TRUE, levls = c(0:10, 15, 20),
    varName = "net")
gof0.gd_5 <- sienaGOF(model_5_effects, GeodesicDistribution, cumulative = FALSE, verbose = FALSE, join = TRUE, varName = "net")
```

### Describe Effects 

```{r GOF Graphs Model 5}
plot(gofi0_5, main = "") # still a tragedy bb
plot(gofo0_5, main = "") #horrible
plot(gof0.gd_5, main = "") #bad (aka: need structural effects)
```

# Models With Structrural Effects {.tabset}

## Race

```{r strctrl white passing}
myeff_1 <- getEffects(the_data)
#view(myeff_1)
myeff_1 <- includeEffects(myeff_1, sameX, interaction1 = "white_pass_covar")
myeff_1 <- includeEffects(myeff_1, egoX, altX, interaction1 ="white_pass_covar")
myeff_1 <- includeEffects(myeff_1, transTrip)
myeff_1 <- includeEffects(myeff_1, transTies)

myAlgorithm <- sienaAlgorithmCreate(projname = "preliminary")
mod_1_effects <- siena07(myAlgorithm, data = the_data, effects = myeff_1, returnDeps = TRUE)
mod_1_effects # 
#write_rds(mod_1_effects, "mod_1_structural_eff_incl_pt2.rds")

```

## Ethnicity

```{r strctrl ethnicity}
myeff_2 <- getEffects(the_data)
#view(myeff_2)
myeff_2 <- includeEffects(myeff_2, sameX, interaction1 = "country_covar")
myeff_2 <- includeEffects(myeff_2, egoX, altX, interaction1 ="country_covar")
myeff_2 <- includeEffects(myeff_2, transTrip)
myeff_2 <- includeEffects(myeff_2, transTies)

myAlgorithm <- sienaAlgorithmCreate(projname = "preliminary")
mod_2_effects <- siena07(myAlgorithm, data = the_data, effects = myeff_2, returnDeps = TRUE)
mod_2_effects # 
#write_rds(mod_2_effects, "mod_2_structural_eff_incl.rds")
```

## Gender 

```{r strctrl gender}
myeff_3 <- getEffects(the_data)
#view(myeff_1)
myeff_3 <- includeEffects(myeff_3, sameX, interaction1 = "gender_covar")
myeff_3 <- includeEffects(myeff_3, egoX, altX, interaction1 ="gender_covar")
myeff_3 <- includeEffects(myeff_3, transTrip)
myeff_3 <- includeEffects(myeff_3, transTies)

myAlgorithm <- sienaAlgorithmCreate(projname = "preliminary")
mod_3_effects <- siena07(myAlgorithm, data = the_data, effects = myeff_3, returnDeps = TRUE)
mod_3_effects # 
#write_rds(mod_3_effects, "mod_3_structural_eff_incl.rds")

```

## Race/Gender
```{r strctrl race gender}
myeff_4 <- includeEffects(myeff_1, sameX, interaction1 = "gender_covar")
myeff_4 <- includeEffects(myeff_4, egoX, altX, interaction1 ="gender_covar")

myAlgorithm <- sienaAlgorithmCreate(projname = "preliminary")
mod_4_effects <- siena07(myAlgorithm, data = the_data, effects = myeff_4, returnDeps = TRUE)
mod_4_effects # 
#write_rds(mod_4_effects, "mod_4_structural_eff_incl.rds")

```

## Ethnicity/Gender

```{r strctrl ethnicity gender}
myeff_5 <- includeEffects(myeff_2, sameX, interaction1 = "gender_covar")
myeff_5 <- includeEffects(myeff_5, egoX, altX, interaction1 ="gender_covar")

myAlgorithm <- sienaAlgorithmCreate(projname = "preliminary")
mod_5_effects <- siena07(myAlgorithm, data = the_data, effects = myeff_5, returnDeps = TRUE)
mod_5_effects # 
#write_rds(mod_5_effects, "mod_5_structural_eff_incl.rds")

```

# GOF Structural Effects {.tabset}
## M1
### Create Effects
```{r GOF functions Model 1 str, message = F, warning=F}
model_1_effects <- read_rds("mod_1_structural_eff_incl_pt2.rds")
gofi0_1 <- sienaGOF(model_1_effects, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "net")
gofo0_1 <- sienaGOF(model_1_effects, OutdegreeDistribution, verbose = FALSE, join = TRUE, levls = c(0:10, 15, 20),
    varName = "net")
gof0.gd_1 <- sienaGOF(model_1_effects, GeodesicDistribution, cumulative = FALSE, verbose = FALSE, join = TRUE, varName = "net")
```

### Describe Effects
```{r GOF Graphs Model 1 str}
plot(gofi0_1, main = "")
plot(gofo0_1, main = "")
plot(gof0.gd_1, main = "") #so these don't work, we need to adjust! (like, with other models? so look at 4)
```

## Model 2

### Create Effects
```{r GOF functions Model 2 str}
model_2_effects <- read_rds("mod_2_structural_eff_incl.rds")
gofi0_2 <- sienaGOF(model_2_effects, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "net")
gofo0_2 <- sienaGOF(model_2_effects, OutdegreeDistribution, verbose = FALSE, join = TRUE, levls = c(0:10, 15, 20),
    varName = "net")
gof0.gd_2 <- sienaGOF(model_2_effects, GeodesicDistribution, cumulative = FALSE, verbose = FALSE, join = TRUE, varName = "net")
```

### Describe Effects
```{r GOF Graphs Model 2 str}
plot(gofi0_2, main = "") # still a tragedy bb
plot(gofo0_2, main = "") #horrible
plot(gof0.gd_2, main = "") #bad (aka: need structural effects)
```


## Model 3

### Create Effects
```{r GOF functions Model 3 str}
model_3_effects <- read_rds("mod_3_structural_eff_incl.rds")
gofi0_3 <- sienaGOF(model_3_effects, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "net")
gofo0_3 <- sienaGOF(model_3_effects, OutdegreeDistribution, verbose = FALSE, join = TRUE, levls = c(0:10, 15, 20),
    varName = "net")
gof0.gd_3 <- sienaGOF(model_3_effects, GeodesicDistribution, cumulative = FALSE, verbose = FALSE, join = TRUE, varName = "net")
```

### Describe Effects
```{r GOF Graphs Model 3 str}
plot(gofi0_3, main = "") # still a tragedy bb
plot(gofo0_3, main = "") #horrible
plot(gof0.gd_3, main = "") #bad (aka: need structural effects)
```

## Model 4

### Create Effects
```{r GOF functions Model 4 str}
model_4_effects <- read_rds("mod_4_structural_eff_incl.rds")
gofi0_4 <- sienaGOF(model_4_effects, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "net")
gofo0_4 <- sienaGOF(model_4_effects, OutdegreeDistribution, verbose = FALSE, join = TRUE, levls = c(0:10, 15, 20),
    varName = "net")
gof0.gd_4 <- sienaGOF(model_4_effects, GeodesicDistribution, cumulative = FALSE, verbose = FALSE, join = TRUE, varName = "net")
```

### Describe Effects
```{r GOF Graphs Model 4 str}
plot(gofi0_4, main = "") # still a tragedy bb
plot(gofo0_4, main = "") #horrible
plot(gof0.gd_4, main = "") #bad (aka: need structural effects)
```

## Model 5

### Create Effects
```{r GOF functions Model 5 str}
model_5_effects <- read_rds("mod_5_structural_eff_incl.rds")
gofi0_5 <- sienaGOF(model_5_effects, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "net")
gofo0_5 <- sienaGOF(model_5_effects, OutdegreeDistribution, verbose = FALSE, join = TRUE, levls = c(0:10, 15, 20),
    varName = "net")
gof0.gd_5 <- sienaGOF(model_5_effects, GeodesicDistribution, cumulative = FALSE, verbose = FALSE, join = TRUE, varName = "net")
```

### Describe Effects 

```{r GOF Graphs Model 5 str}
plot(gofi0_5, main = "") # still a tragedy bb
plot(gofo0_5, main = "") #horrible
plot(gof0.gd_5, main = "") #bad (aka: need structural effects)
```
