---
title: "The Dataset"
output: html_document
date: "2024-10-16"
---
```{r setup packages, include=FALSE}
require(tidyverse)
require(igraph)
require(shiny)
require(rmdformats)
require(prettydoc)
require(hrbrthemes)
require(tint)
require(tufte)
require(psych)
library(data.table)  
library(xml2)
library(rvest)
library(XML)
```

```{r setup data, include=FALSE}
load("/Users/hannah/Desktop/Season 7, Part 1/Social Networks/labjournal/scholars_20240924.rda")
scholars <- x
rm(x)
```

# The Parameters and Variables (Some of them, at Least)

\

**[data]**: our scholars file\
**university**: Character vector with names of universities. We have several universities in the Netherlands. See above for relevant names.\
**discipline**: Character vector, either sociology or political science or both.\
**waves**: a list of numeric vectors with start and end year of wave.\
**type:**\
**ties**
- *first*: directed: first author sending to others\
- *last*: directed: last author sending to others\
- *all*: undirected: ties between all authors\
**Output:** - a list\
- *nets*: array of nomination networks.\
- *data*: sample of data (scholars)\

```{r function dataset, include = F, echo = F, message=FALSE, warning=F}
#so this is all we are doin'
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}

```

# Look at the Data 

Now we can use the data, that we have cleaned up a bit, to make a network-plot, and see what we're dealing with.\

We'll start easy though, to get a base-level intuitive understanding of what the data look like -- and for that, we're using only the data of Radboud's sociology department in our first wave. 
\

## RU Sociology Graph 
```{r matrix 1, fig.dim = c(8, 6)}
#save the output of your function
matrix_ru_soc <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = "sociology", 
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))

mrs_graph1 <- igraph::graph_from_adjacency_matrix(
  matrix_ru_soc$nets[1,,], #for this example I take the first wave of data. (thus I select the array of networks and take the first matrix)
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL
)

plot(mrs_graph1,
  vertex.label= NA,
  edge.width = 0.2,
  edge.arrow.size =0.2)
```

Then make it a bit more complicated, and add a new discipline (political science), and another university. 

## RU & UvT Sociology and Political Science Graph 

```{r,  fig.dim = c(8, 6)}
matrix_ru_socpol <- fcolnet(data = scholars, 
                university = c("RU", "UvT"), 
                discipline = c("sociology", "political science"), 
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("all"))

mrsp <- igraph::graph_from_adjacency_matrix(
  matrix_ru_socpol$nets[2,,], #now, I take the second wave
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL
)

plot(mrsp,
     vertex.label = NA,
     vertex.size = .5,
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)
```

... and see if we can create some differentiations in the plot to see how different characteristics interact (discipline, university)

```{r now some more,  fig.dim = c(8, 6)}
#Let us find ego characteristics. 
matrix_ru_socpol_df <- matrix_ru_socpol$data
mrusp_df_ego <- do.call(rbind.data.frame, matrix_ru_socpol_df$demographics)

plot(mrsp,
  vertex.color = ifelse(mrusp_df_ego$discipline.24 == "sociology", "red", "blue"), #now, I can use actor attributes for plotting. 
  vertex.label = NA,
  edge.width = 0.2,
  edge.arrow.size =0.2,
  vertex.size = 5)
```

## My Four Departments -- Graph

```{r}
matrix_my_people <- fcolnet(data = scholars, 
                university = c("RU", "UU", "UvA", "EUR"), 
                discipline = "sociology",
                waves = list(c(2015, 2019), c(2020, 2024)), 
                type = c("first"))

graph_my_people <- igraph::graph_from_adjacency_matrix(
  matrix_ru_socpol$nets[1,,], #first wave
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL
)

plot(mrsp,
     vertex.label = NA,
     vertex.size = 10,
     edge.curved = 0.2,
     edge.arrow.size = 0.1)
```
... and add some ego-level characteristics in this graph!!

```{r}
matrix_my_people_df <- matrix_my_people$data
df_ego_my_people <- do.call(rbind.data.frame, matrix_my_people_df$demographics)

plot(graph_my_people,
  vertex.color = ifelse(df_ego_my_people$Universiteit.24 == "UU", "red", "blue"), #now, I can use actor attributes for plotting. 
  vertex.label = NA,
  edge.width = 0.2,
  edge.arrow.size =0.2,
  vertex.size = 10)

```

```{r write df, include = F, echo = F}
#write_rds(df_ego_my_people, "demographics_my_people.rds")
```

