---
title: "Data Visualisation"
output: html_document
date: "2024-10-08"
---

```{r setup a, include=FALSE}
require(tidyverse)
require(igraph)
require(shiny)
require(RSiena)
require(rmdformats)
require(prettydoc)
require(hrbrthemes)
require(tint)
require(tufte)
require(psych)
```

```{r jochem functions, include = F, echo = F}
fpackage.check <- function(packages) {
  lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
  ifelse(!dir.exists("data"), dir.create("data"), FALSE)
  ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
  if (is.null(file))
    file = deparse(substitute(x))
  datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
  totalname <- paste(location, datename, file, ".rda", sep = "")
  save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
  load(filename)
  get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
  knitr::kable(x, digits = 2, "html", ...) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
    kableExtra::scroll_box(width = "100%", height = "300px")
}

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology", affiliation = "RU",
                      waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {
  
  publications <- list_publications %>%
    bind_rows() %>%
    distinct(title, .keep_all = TRUE)
  
  df_scholars %>%
    filter(affil1 == affiliation | affil2 == affiliation) %>%
    filter(discipline == discip) -> df_sel
  
  networklist <- list()
  for (wave in 1:length(waves)) {
    networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
  }
  
  publicationlist <- list()
  for (wave in 1:length(waves)) {
    publicationlist[[wave]] <- publications %>%
      filter(gs_id %in% df_sel$gs_id) %>%
      filter(year %in% waves[[wave]]) %>%
      select(author) %>%
      lapply(str_split, pattern = ",")
  }
  
  publicationlist2 <- list()
  for (wave in 1:length(waves)) {
    publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
      # lowercase
      lapply(tolower) %>%
      # Removing diacritics
      lapply(stri_trans_general, id = "latin-ascii") %>%
      # only last name
      lapply(word, start = -1, sep = " ") %>%
      # only last last name
      lapply(word, start = -1, sep = "-")
  }
  
  for (wave in 1:length(waves)) {
    # let us remove all publications with only one author
    remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
    publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
  }
  
  for (wave in 1:length(waves)) {
    pubs <- publicationlist2[[wave]]
    for (ego in 1:nrow(df_sel)) {
      # which ego?
      lastname_ego <- df_sel$lastname[ego]
      # for all publications
      for (pub in 1:length(pubs)) {
        # only continue if ego is author of pub
        if (lastname_ego %in% pubs[[pub]]) {
          aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
          # only continue if ego is first author of pub
          if (aut_pot == 1) {
            # check all alters/co-authors
            for (alter in 1:nrow(df_sel)) {
              # which alter
              lastname_alter <- df_sel$lastname[alter]
              if (lastname_alter %in% pubs[[pub]]) {
                networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
              }
            }
          }
        }
      }
    }
  }
  return(list(df = df_sel, network = networklist))
}

fcolnet <- function(data = scholars, university = "RU", discipline = "sociology", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}

```

----
```{r setup 2, include=FALSE}
load("/Users/hannah/Desktop/Season 7, Part 1/Social Networks/labjournal/scholars_20240924.rda")
scholars <- x
rm(x)
```

```{r copied from jochem 1, results = F, include = F, echo = F}
#copied from jochem
# (sociology, RU)
demographics <- do.call(rbind.data.frame, scholars$demographics)
demographics <- demographics %>% # who's who
    mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
        is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
        ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
        is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

sample <- which((demographics$Universiteit1.22 == "RU" | demographics$Universiteit2.22 == "RU" | demographics$Universiteit1.24 ==
    "RU" | demographics$Universiteit2.24 == "RU") & (demographics$discipline.22 == "sociology" | demographics$discipline.24 ==
    "sociology")) #sociologists at ru

demographics_soc <- demographics[sample, ]
scholars_sel <- lapply(scholars, "[", sample)
```

```{r copied from jochem 2, results = F,include = F, echo = F}
ids <- demographics_soc$au_id #openalex ids
wave2 <- wave1 <- matrix(0, nrow = length(ids), ncol = length(ids), dimnames = list(ids, ids))
#matrix for all ids
```

```{r copied from jochem 3, results = F, include = F, echo = F}
works <- scholars_sel$work
works_id <- unlist(lapply(works, function(l) l$id)) #id for works
works_author <- unlist(lapply(works, function(l) l$author), recursive = FALSE) #authors of the work
works_year <- unlist(lapply(works, function(l) l$publication_year), recursive = FALSE) #year
df_works <- tibble(works_id, works_author, works_year) #id-work, authors, year
dups <- which(duplicated(works_id))
# why are there some many papers of Batenburg duplicates

df_works <- df_works[-dups, ]
df_works_newer <- df_works[df_works$works_year > 2014, ]
df_works_w2 <- df_works[df_works$works_year > 2019, ] 
```

```{r jochem deed dit, include = F, results = F, echo = F}
# Chunk 4

## If we wanted to, we could

#first to others
for (i in 1:nrow(df_works_w2)) {
    ego <- df_works_w2$works_author[i][[1]]$au_id[1]
    alters <- df_works_w2$works_author[i][[1]]$au_id[-1]

    if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
        wave2[which(ids %in% ego), which(ids %in% alters)] <- 1
    }
} 
#for each row, of dataframes of work, do egos (first auth) and alters (others) on matrix (and then: are they the same or not)

#last to others
for (i in 1:nrow(df_works_w2)) {
    ego <- rev(df_works_w2$works_author[i][[1]]$au_id)[1]
    alters <- rev(df_works_w2$works_author[i][[1]]$au_id)[-1]

    if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
        wave2[which(ids %in% ego), which(ids %in% alters)] <- 1
    }
} 
#for each row, of dataframes of work, do egos (last auth) and alters (others) on matrix (and then: are they the same or not)

#undirected
for (i in 1:nrow(df_works_w2)) {
    egos <- df_works_w2$works_author[i][[1]]$au_id

    if (sum(ids %in% egos) > 0) {
        wave2[which(ids %in% egos), which(ids %in% egos)] <- 1
    }
}
#for each row, of dataframes of work, do egos (all authors) on matrix (and then: are they the same or not)
```


# What Do Our Data Look Like 

Before we can decide how to deal with subsetting our data, we first want to establish what kind of universities we are dealing with -- which Dutch universities are included, and are there any universities that *are not* Dutch. 
\
```{r universities, results=F}
universities_1_2022 <- unique(demographics$Universiteit1.22)
universities_2_2022 <- unique(demographics$Universiteit2.22)
universities_1_2024 <- unique(demographics$Universiteit1.24)
universities_2_2024 <- unique(demographics$Universiteit2.24)

universities_1_2022 
universities_2_2022
universities_1_2024
universities_2_2024

#"UU", "RUG", "UvA", "EUR", "Leiden", "Bocconi University", RU", "Universita degli studi di Milano", "VU", "UvT", "Universiteit Gent", "Universiteit Stockholm", "University Linköping", "TU Delft", "University of Turku", "Frankfurt School of Finance & Management", "University of Lausanne", "UvH", "Boston University", "Trento University", "Uni Gothenburg", "Uni Berlijn", "Uni of Victoria", "Uni of Lausanne", "Shadona uni", "University of Cologne","Shandong uni"   
```
\

The Universities\


|Abbreviation | Name and City | Country |
|:---------:|:---------:|:---------:|
|"UU"| Utrecht University | the Netherlands |
|"RUG"| Rijksuniversiteit Groningen | the Netherlands|
|"UvA"| Universiteit van Amsterdam | the Netherlands|
|"EUR"| Erasmus Universiteit Rotterdam | the Netherlands |
|"Leiden"| Leiden University | the Netherlands |
|"Bocconi University"| Bocconi University, Milan | Italy |
| RU" | Radboud University | the Netherlands |
|"Universita degli studi di Milano"| Universita degli studi di Milano | Italy |
|"VU"| Vrije Universiteit, Amsterdam | the Netherlands |
|"UvT"| Tilburg University | the Netherlands |
|"Universiteit Gent"| University Gent| Belgium |
|"Universiteit Stockholm"| University of Stockholm | Sweden |
|"University Linköping"| University of Linköping | Sweden |
|"TU Delft"| TU Delft | the Netherlands |
|"University of Turku"| University of Turku | Finland |
|"Frankfurt School of Finance & Management"| Frankfurt School of Finance & Management | Germany |
|"University of Lausanne"| University of Lausanne | Switzerland |
|"UvH"| University of Humanistic Studies, Utrecht | the Netherlands |
|"Boston University"| Boston University | the US |
|"Trento University"| Trento Univeristy | Italy |
|"Uni Gothenburg"| University of Gothenburg | Sweden |
|"Uni Berlijn"| University of Berling | Germany |
|"Uni of Victoria"| University of Victoria | Canada |
|"Uni of Lausanne"| University of Lausanne | Switzerland |
|"Shadona uni"| Shadong University | China | 
|"University of Cologne"| University of Cologne | Germany |
|"Shandong uni"| Shadong University | China |
\

Now see if we can put this into a social network plot/graph/whatever. 

```{r}
matrix_all_socpol <- fcolnet(data = scholars, 
                university = c("UU", "RUG", "UvA", "EUR", "Leiden", "RU", 
                               "VU", "UvT", "TU Delft", "UvH"), 
                discipline = c("sociology", "political science"), 
                waves = list(c(2015, 2018), c(2019, 2023))) 

mall_w1 <- igraph::graph_from_adjacency_matrix(
  matrix_all_socpol$nets[1,,],
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL
)
mall_w2 <- igraph::graph_from_adjacency_matrix(
  matrix_all_socpol$nets[2,,],
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL
)

#nicely_chosen <- layout_nicely(mall_w1)
#kk <- layout_with_kk(mall_w1)
#lgl <- layout_with_lgl(mall_w1)
random_1 <- layout_randomly(mall_w1)
random_2 <- layout_randomly(mall_w2)
```

```{r}
matrix_all_socpol_df <- matrix_all_socpol$data
mrusp_df_ego <- do.call(rbind.data.frame, matrix_all_socpol_df$demographics)

plot(mall_w1,
     vertex.color = ifelse(mrusp_df_ego$discipline.22 == "sociology", "#FFC107", "#004D40"),
     vertex.size = 4,  # we'll vertices a bit smaller
     vertex.label = NA,  
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)
plot(mall_w2,
     vertex.color = ifelse(mrusp_df_ego$discipline.24 == "sociology", "#FFC107", "#004D40"),
     vertex.size = 4,  # we'll vertices a bit smaller
     vertex.label = NA,  
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)
```
\
OK, so that worked! -- I think that if I add my gender variables (spoilers! for the next part) to this, I'll also be able to use that to create some fun. 




