---
title: "Week 2"
author: "Hannah Groennou"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: paper
    toc: yes
    toc_float:
      collapsed: true
---

```{r libraries, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(igraph)
require(hrbrthemes)
```

---

# Acquire Matrices

\

## Simple Matrix
First, lets make our own simple matrix, so we can get some feeling with the inner workings of matrices, and how adjustments in matrices relate to different networks and network descriptives.
```{r first matrix}
matrix <- matrix(c(0,0,0,1,
                   1,0,0,0,
                   0,0,1,0,
                   0,0,0,0), nrow = 4, ncol = 4)
graph <- graph_from_adjacency_matrix(matrix)
matrix_graph <- triad_census(graph)
matrix_graph
```
\


## RSiena *s501* and *s502*
Then we download a few pre-installed matrices, and subset a smaller sample to make checking the results easier.

```{r all matrix}
t1 <- s501 #matrix at t1
t2 <- s502 #matrix at t2
dim(s501) #dimensions full matrix
small_t1 <- s501[c(1:10), c(1:10)] #sample matrix t1
small_t2 <- s502[c(1:10), c(1:10)] #sample matrix t2
small_graph_t1 <- graph_from_adjacency_matrix(small_t1)
small_graph_t2 <- graph_from_adjacency_matrix(small_t2)
plot(small_graph_t1) #graph of sample matrix t1
plot(small_graph_t2) #graph of sample matrix t2, some broken relations I see

```

---

# Descriptives Small Sample

To better understand our sample, we look at some descriptives to understand some core aspects of the matrix, and the network.

\


## Density
Density concerns the percentage of possible ties realised in the network. In other words, it is calculated by taking the amount of ties present in the sample, and dividing this by the total possible amount of ties in a --- in this case --- network with 10 nodes.\
```{r small density}
density_t1 <- edge_density(small_graph_t1)
density_t1 #so that's approximately 11%
```

\

## Centrality 
Then we look at the centrality of each node, which concerns the number of edges each one has. The more edges a node has, the more central they are. As such, this essentially measures how important or well connected each vertix is.\
\
Centrality scores can either be calculated to yield raw, or normalised measures. Raw measures are particular to networks or samples, and as such cannot be used to compare across samples, but only to investigate within one. Normalised measures, on the other hand, allow one to compare across networks or samples --- which is often more useful in research.\
```{r small degree centrality}
degree_t1 <- igraph::degree(small_graph_t1)#don't care about directions 
degree_t1 #not normalised, it seems 
degree_t1_norm_a <- igraph::degree(small_graph_t1, normalized = T)
degree_t1_norm_a #normalised using only sample? 

#let's see how it differs if we calculate degree according to the formula given by Jochem
normalised_cent_t1 <- (degree_t1 - min(degree_t1))/(max(degree_t1)) - min(degree_t1)
normalised_cent_t1 #significanly different, it appears

```

\

## Closeness Centrality
Closeness centrality of a vertix, or node, is the average distance between the node of concern, and the other nodes it is connected with in the sample. It is calculated by taking the total number of nodes in a vertix' network, the sum of distances between the node of concern, and subsequently dividing the former by the latter. 

```{r small closeness centrality}
small_t1 <- s501[c(1:10), c(1:10)] #sample matrix t1
small_graph_t1 <- graph_from_adjacency_matrix(small_t1)
closeness_t1 <- igraph::closeness(small_graph_t1)
closeness_t1
```
\
*n.b.* the 'NaN' in the output refers to the fact that the vertices in question lack any ties, and thus result in an error when attempting to calculate the closeness.\
\


## Betweenness Centrality

Betweenness centrality is essentially the number of shortest paths between vertix *a* and vertix *b* while going through vertix *c*. In order to be able to compare the measure between samples, a normalised version can be calculated by first subtracting from the *betweenness centrality* of the certain vector from the minimum score in the sample, and dividing it by the range of *betweenness centrality* in the entire sample.

```{r small betweenness centrality}
between_t1 <- igraph::betweenness(small_graph_t1, directed = T, normalized = F)
between_t1 #zeroes
between_t1_norm <- igraph::betweenness(small_graph_t1, directed = T, normalized = T)
between_t1_norm #normalisation does not matter, because the zeroes 
```
\
Due to the structure of our sample, where the largest closed networks consist of only 3 nodes, betweenness centrality does not yield a meaningful measure because it is not present; as demonstrated by the output.\
\

## Clustering

Clustering refers to the amount of 'clumping' one might see in a network. It is quantified as the proportion of actual- compared to possible- ties of a specific vertix. The level of clustering is therefore a characteristic of an ego-network and --- when it concerns undirected ties --- the same as a transitivity index.
```{r small clustering}
transitivity_t1 <- igraph::transitivity(small_graph_t1, type = "local")
transitivity_t1
```


