---
title: "Week 3: Webscraping"
output: html_document
date: "2024-09-18"
---

```{r setup, include=FALSE}
#install.packages("DataCombine")
library(DataCombine)
library(tidyverse)
library(data.table)
library(httr)
library(xml2)
library(devtools)
library(scholar)
```

# Webscraping For Dummies: The Staff-Info Edition

So, we are going to do some good-good webscraping for dummies!

```{r first part, echo=TRUE, fig.height=4}
#first, we are going to 'scrape' something from the Radboud website - the members of the sociology department
soc_staff_ru <- read_html("http://web.archive.org/web/20230528153336/https://www.ru.nl/sociology/research/staff/")
str(soc_staff_ru)
soc_staff_ru <- soc_staff_ru %>%
  rvest::html_element("body") %>%
  xml2::xml_find_all("//td") %>%
  rvest::html_text()
#ok maybe everything from the same packages
soc_staff_ru <- read_html("http://web.archive.org/web/20230528153336/https://www.ru.nl/sociology/research/staff/")
soc_staff_ru <- soc_staff_ru %>%
  rvest::html_element("body") %>%
  rvest::html_elements("td") %>%
  rvest::html_text()

view(soc_staff_ru)
```

```{r better, echo=TRUE, fig.height=4}
#now Better
soc_staff <- read_html("http://web.archive.org/web/20230528153336/https://www.ru.nl/sociology/research/staff/")
soc_staff <- soc_staff %>%
  rvest::html_element("body")
soc_staff <- soc_staff %>%
  rvest::html_table()
view(soc_staff)  
```

Then we go to another department: Leiden's Political Science Department
```{r staff leiden, include = F, echo = T}
#find the best link to give us all the staff of interest on one webpage 
pol_staff_leiden_b <- read_html("https://www.universiteitleiden.nl/en/social-behavioural-sciences/political-science/staff#tab-1") #page 
#str(pol_staff_leiden_b)
#extract the information - shortest way to go 
pol_staff_leiden_all <- pol_staff_leiden_b %>% #names of staff
  rvest::html_elements("ul") %>%
  rvest::html_elements("li") %>%
  rvest::html_elements("a") %>%
  rvest::html_elements("div") %>%
  rvest::html_text()
#extract the information - longer way to go 
pol_staff_leiden_b <- read_html("https://www.universiteitleiden.nl/en/social-behavioural-sciences/political-science/staff#tab-1") #page 
pol_staff_leiden_names <- pol_staff_leiden_b %>% #names of staff
  rvest::html_elements("ul") %>%
  rvest::html_elements("li") %>%
  rvest::html_elements("a") %>%
  rvest::html_elements("div") %>%
  rvest::html_elements("strong") %>%
  rvest::html_text()
#now the positions of the staff
pol_staff_leiden_b <- read_html("https://www.universiteitleiden.nl/en/social-behavioural-sciences/political-science/staff#tab-1") #page 
pol_staff_leiden_position <- pol_staff_leiden_b %>% #position of staff
  rvest::html_elements("ul") %>%
  rvest::html_elements("li") %>%
  rvest::html_elements("a") %>%
  rvest::html_elements("div") %>%
  rvest::html_elements("span") %>%
  rvest::html_text()
#see where the missing information is
view(pol_staff_leiden_position)
view(pol_staff_leiden_names)
#create NA value
newrow <- "NA"
index <- 76
#make dataframe
pol_staff_leiden_position <- as.data.frame(pol_staff_leiden_position)
#add the NA to the dataframe
pol_staff_leiden_position <- InsertRow(pol_staff_leiden_position, newrow, index)
view(pol_staff_leiden_names)
view(pol_staff_leiden_position)
view(pol_staff_leiden_all)
```

# But What If I Want to Know More?

Well, sure we have some fun info about the names and positions of the staff. But how can we use this to *do* something with the information./
Let me tell you how you can use this! By using the names as starting points for setting up collaboration networks./
/
Because we have these names, we can now use **webscraping** to find additional information associated with these names -- such as the papers they have published, and with whom they have been published./
We do this by going to the OpenAlex API, and extracting our information of choice accordingly.

## Get things started 

```{r try open alex first part, echo=TRUE, fig.height=4}
#install.packages("jsonlite") #making json objects
#install.packages("openalexR") #OpenAlex for R
library(openalexR)
library(jsonlite)

options(openalexR.mailto = "hannah@groennou.nl")  #my own email bb
url <- "https://api.openalex.org/authors?search=Jochem Tolsma"

jt_json <- fromJSON("https://api.openalex.org/authors?search=Jochem+Tolsma", simplifyVector = FALSE)
glimpse(jt_json, max.level = 1)
#jt_json[["results"]][[1]][["display_name"]]
#df_jt <- jt_json %>%
#    .$results %>%
#  .[[1]] %>%
#    discard(is_empty)
```

## Find some Jochem info 
```{r Jochem open alex, echo=TRUE, fig.height=4}
df_jochem <- oa_fetch(entity = "author", search = "Jochem Tolsma")
view(df_jochem) #we've lost his secondary information --- so if you want to decide on wrapper or direct interaction w/the API

df_jochem_papers <- oa_fetch(entity = "works", author.id = df_jochem$id)
str(df_jochem_papers)
view(df_jochem_papers)
```

## Use the Wrapper 

Because we are (weak), we are using the wrapper (becasue it's easier)
\
Get the people first 
```{r filters, echo=TRUE, fig.height=4}
options(openalexR.mailto = "hannah.groennou@ru.nl") #indicate you're you

ru_id <- institutions <- oa_fetch(entity = "institutions", search = "Radboud University") #get the id of the radboud university 
ru_id

uva_id <- oa_fetch(entity = "institutions", search = "Universiteit van Amsterdam") #get uva's id 
uva_id

jochem_tolsma <- oa_fetch(entity = "author", search = "Jochem Tolsma", affiliations.institution.id = "https://openalex.org/I145872427") #get jochem info

tom_vd_meer  <- oa_fetch(entity = "author", search = "Tom van der Meer", affiliations.institution.id = uva_id$id) #Toni is in there, but let's ignore him because we are only using Tom-1  (get tom info)
tom_vd_meer <- tom_vd_meer[1,] #get the RIGHT tom 

maurice_gesthuizen <- oa_fetch(entity = "author", search = "Maurice Gesthuizen", affiliations.institution.id = "https://openalex.org/I145872427") #maurice

michael_savelkoul <- oa_fetch(entity = "author", search = "Michael Savelkoul", affiliations.institution.id = "https://openalex.org/I145872427") #michael
```

First make separate objects for all their works 
\
```{r all papers, echo=TRUE, fig.height=4}
jochem_tolsma_papers <- oa_fetch(entity = "works", author.id = jochem_tolsma$id) #jochem
tom_vd_meer_papers <- oa_fetch(entity = "works", author.id = tom_vd_meer$id) #tom
maurice_gesthuizen_papers <- oa_fetch(entity = "works", author.id = maurice_gesthuizen$id) #maurice
michael_savelkoul_papers <- oa_fetch(entity = "works", author.id = michael_savelkoul$id) #michael
```

Then create one big object, in which we collect all four people's work (all of 'em)
\
```{r make all works together, echo=TRUE, fig.height=4}
all_works <- rbind(jochem_tolsma_papers, tom_vd_meer_papers)
all_works <- rbind(all_works, maurice_gesthuizen_papers)
all_works <- rbind(all_works, michael_savelkoul_papers)
```

Then we create an adjacency matrix -- first by filtering out the data from only the relevant years (2019 - 2023)

```{r filter out years, echo=TRUE, fig.height=4}
all_works <- all_works %>% #filter out years 
  filter(
    publication_year >= 2019) %>%
  filter(
    publication_year <= 2023)
```

Then, we put all the information in a matrix -- to create an adjacency matrix to assess patterns of collaboration
\
```{r make matrix, echo=TRUE, fig.height=4}
adjacency_matrix <- matrix(0, nrow = 4, ncol = 4) #first we make an empty matrix 
  
dimnames(adjacency_matrix) #then we ask for the names of the rows and columns 
rownames(adjacency_matrix) <- c(jochem_tolsma$id, tom_vd_meer$id, maurice_gesthuizen$id, 
                                michael_savelkoul$id) #name the rows after id's 
colnames(adjacency_matrix) <- c(jochem_tolsma$id, tom_vd_meer$id, maurice_gesthuizen$id, 
                                michael_savelkoul$id) #name the columns after id's
adjacency_matrix #now what does the matrix look like?

view(all_works) #so what's the dataset look like again 
ego <- all_works[1,]$author[[1]]$au_id[1] #select the first author of the first work (ego)
view(ego)
alters <-  all_works[1,]$author[[1]]$au_id[-1] #select the non-first author of the first work (alter)
view(alters)

#they refuse rn. see later why 
#adjacency_matrix[ego, alters] <- 1 
#view(adjacency_matrix)
#adjacency_matrix[ego, alters] <- 1 #once more
```

