---
title: "Scraping Some Data"
output: html_document
date: "2024-09-29"
---

```{r libraries a, include = F, echo = F}
#rm(list = ls())
require(tidyverse)
require(igraph)
require(shiny)
require(RSiena)
require(rmdformats)
require(prettydoc)
require(hrbrthemes)
require(tint)
require(tufte)
require(psych)
library(data.table)  
library(xml2)
library(rvest)
library(xml2)
library(XML)
```
```{r setup a, include = F, echo = F}
load("/Users/hannah/Desktop/Season 7, Part 1/Social Networks/labjournal/scholars_20240924.rda")
scholars <- x
rm(x)
```
```{r copied from jochem 1a, include = F, echo = F}
#copied from jochem
# (sociology, RU)
demographics <- do.call(rbind.data.frame, scholars$demographics)
demographics <- demographics %>% # who's who
    mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
        is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
        ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
        is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

sample <- which((demographics$Universiteit1.22 == "RU" | demographics$Universiteit2.22 == "RU" | demographics$Universiteit1.24 ==
    "RU" | demographics$Universiteit2.24 == "RU") & (demographics$discipline.22 == "sociology" | demographics$discipline.24 ==
    "sociology")) #sociologists at ru

demographics_soc <- demographics[sample, ]
scholars_sel <- lapply(scholars, "[", sample)
```
```{r copied from jochem 2a, include = F, echo = F}
ids <- demographics_soc$au_id #openalex ids
wave2 <- wave1 <- matrix(0, nrow = length(ids), ncol = length(ids), dimnames = list(ids, ids))
#matrix for all ids
```
```{r copied from jochem 3a, include = F, echo = F}
works <- scholars_sel$work
works_id <- unlist(lapply(works, function(l) l$id)) #id for works
works_author <- unlist(lapply(works, function(l) l$author), recursive = FALSE) #authors of the work
works_year <- unlist(lapply(works, function(l) l$publication_year), recursive = FALSE) #year
df_works <- tibble(works_id, works_author, works_year) #id-work, authors, year

dups <- which(duplicated(works_id))
# why are there some many papers of Batenburg duplicates

df_works <- df_works[-dups, ]
df_works_w2 <- df_works[df_works$works_year > 2019, ] #wave 2 only above 2019
```
```{r copied from jochem 5a, include = F, echo = F}
#so this is all we are doin'
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}

```
```{r jochem function, echo = F, include = F}
fcolnet <- function(data = scholars, university = "RU", discipline = "sociology", waves = list(c(2015,
    2018), c(2019, 2023)), type = c("first")) {

    # step 1
    demographics <- do.call(rbind.data.frame, data$demographics)
    demographics <- demographics %>%
        mutate(Universiteit1.22 = replace(Universiteit1.22, is.na(Universiteit1.22), ""), Universiteit2.22 = replace(Universiteit2.22,
            is.na(Universiteit2.22), ""), Universiteit1.24 = replace(Universiteit1.24, is.na(Universiteit1.24),
            ""), Universiteit2.24 = replace(Universiteit2.24, is.na(Universiteit2.24), ""), discipline.22 = replace(discipline.22,
            is.na(discipline.22), ""), discipline.24 = replace(discipline.24, is.na(discipline.24), ""))

    sample <- which((demographics$Universiteit1.22 %in% university | demographics$Universiteit2.22 %in%
        university | demographics$Universiteit1.24 %in% university | demographics$Universiteit2.24 %in%
        university) & (demographics$discipline.22 %in% discipline | demographics$discipline.24 %in% discipline))

    demographics_soc <- demographics[sample, ]
    scholars_sel <- lapply(scholars, "[", sample)

    # step 2
    ids <- demographics_soc$au_id
    nwaves <- length(waves)
    nets <- array(0, dim = c(nwaves, length(ids), length(ids)), dimnames = list(wave = 1:nwaves, ids,
        ids))
    dimnames(nets)

    # step 3
    df_works <- tibble(works_id = unlist(lapply(scholars_sel$work, function(l) l$id)), works_author = unlist(lapply(scholars_sel$work,
        function(l) l$author), recursive = FALSE), works_year = unlist(lapply(scholars_sel$work, function(l) l$publication_year),
        recursive = FALSE))

    df_works <- df_works[!duplicated(df_works), ]

    # step 4
    if (type == "first") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- df_works_w$works_author[i][[1]]$au_id[1]
                alters <- df_works_w$works_author[i][[1]]$au_id[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "last") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                ego <- rev(df_works_w$works_author[i][[1]]$au_id)[1]
                alters <- rev(df_works_w$works_author[i][[1]]$au_id)[-1]
                if (sum(ids %in% ego) > 0 & sum(ids %in% alters) > 0) {
                  nets[j, which(ids %in% ego), which(ids %in% alters)] <- 1
                }
            }
        }
    }

    if (type == "all") {
        for (j in 1:nwaves) {
            df_works_w <- df_works[df_works$works_year >= waves[[j]][1] & df_works$works_year <= waves[[j]][2],
                ]
            for (i in 1:nrow(df_works_w)) {
                egos <- df_works_w$works_author[i][[1]]$au_id
                if (sum(ids %in% egos) > 0) {
                  nets[j, which(ids %in% egos), which(ids %in% egos)] <- 1
                }
            }
        }
    }
    output <- list()
    output$data <- scholars_sel
    output$nets <- nets
    return(output)
}
```
```{r something, include = F, echo = F}
#save the output of your function
matrix_ru_soc <- fcolnet(data = scholars, 
                university = "RU", 
                discipline = "sociology", 
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))

mrs_graph1 <- igraph::graph_from_adjacency_matrix(
  matrix_ru_soc$nets[1,,], #for this example I take the first wave of data. (thus I select the array of networks and take the first matrix)
  mode = c("directed"),
  weighted = NULL,
  diag = FALSE,
  add.colnames = NULL,
  add.rownames = NULL
)

plot(mrs_graph1,
  vertex.label= NA,
  edge.width = 0.2,
  edge.arrow.size =0.2)
```
```{r packages, include=F, echo=F}
# first we get a library, to install something
library(tidyverse)
library(foreign)
library(haven)
library(labelled)
library(devtools)
#devtools::install_github("kalimu/genderizeR")
library(genderizeR)
library(httr)
library(xml2)
library(reshape2)
library(stringr)
```
```{r custom packages, include = F, echo = F}
#'own' packages----
fpackage.check <- function(packages) {
  lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
  ifelse(!dir.exists("data"), dir.create("data"), FALSE)
  ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
  if (is.null(file))
    file = deparse(substitute(x))
  datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
  totalname <- paste(location, datename, file, ".rda", sep = "")
  save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
  load(filename)
  get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
  knitr::kable(x, digits = 2, "html", ...) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
    kableExtra::scroll_box(width = "100%", height = "300px")
}

colorize <- function(x, color) {
  sprintf("<span style='color: %s;'>%s</span>", color, x)
}

```

# The Data 

For our project, we're working with a dataset containing all works written by all academic staff at all Dutch universities in 2022 and 2024. Before we can do anything with these data, we're going to visualise it to get some quick info for our eyeballs to process. 

```{r graphs first time}
matrix_ru_soc <- fcolnet(data = scholars, 
                university = c("UU", "RUG", "UvA", "EUR", "Leiden"),
                discipline = "sociology",
                waves = list(c(2015, 2018), c(2019, 2023)), 
                type = c("first"))
#demographics <- as.data.frame(demographics)
#demographics$dep_col <- if_else(demographics$discipline.24 == "political science", "#D81B60", "#1E88E5")
#net1_g <- graph_from_adjacency_matrix(matrix_ru_soc, mode = c("undirected"))
#plot(net1_g, #vertex.color = demographics$dep_col,
#  vertex.size = 4,  
# vertex.label = NA,
#  edge.curved = 0.2, 
#  edge.arrow.size = 0.1) #ok so that's uneventful
```
Ok, something absolutely and categorically failed to work, but that's for future me -- because time is a thing and there is a max. 

\

In order to facilitate the answering of the questions we have formulated, some data-wrangling has to be done first. We're going to use our data on the demographics of our sample of scientists to assess, first, their gender, and then their ethnicity. 

```{r data}
#demographics <- read_csv("demographics.csv") #load the demographics 
```

# Data Transformation -- Get Some Gender 

Having loaded the data into out Global Environment, we can now extract the first thing we are going to need in order to be able to ascertain each individual's gender, based on their first names. 

```{r split names}
head(demographics) #so what does our demographics data look like 
#then we create a vector of all names
names_NL <- demographics$Naam
#and split names in 'first' and 'last'
names_sep <- stringr::str_split_fixed(names_NL, "\\s", 2) #note: this is not perfect, the dutch have some fascinating 'tussenvoegsels', but that's something we're going to have to deal with one's we get to the ethnicity thing. 
names_sep <- as.data.frame(names_sep) #then we create a df with the seperated names t
au_id <- demographics$au_id #we'll also extract their openalex id to add to the df
names <- cbind(names_sep, au_id) #and col.bind this openalex id with the the seperate names 
view(names) #look if all's gone well

names <- names %>% #then, to make things a bit easier to understand later on, we rename the variables
  rename(
    first_name = V1,
    last_name = V2
  )
#and then we change the value lables, so it's even easier to understand 
var_label(names$first_name) <- "A person's first name" 
var_label(names$last_name) <- "A person's last name"
var_label(names$au_id) <- "One's OpenAlex ID"
view(names)
```

Having prepped the data, we are now going to use a wrapper for the GenderizeR api, to assess the gender of all first names. 

```{r genderize names}
library(genderizeR) #load the package
#only small chunks of names work, large one's don't 
gendered_names_1 <- genderizeAPI(names$first_name[c(1:10)], country = NULL, ssl.verifypeer = T ) #we can only ask for 100 names a day, so we're doing the first 100
gendered_names_2 <- genderizeAPI(names$first_name[c(11:20)], country = NULL, ssl.verifypeer = T )
gendered_names_3 <- genderizeAPI(names$first_name[c(21:30)], country = NULL, ssl.verifypeer = T )
gendered_names_4 <- genderizeAPI(names$first_name[c(31:40)], country = NULL, ssl.verifypeer = T )
gendered_names_5 <- genderizeAPI(names$first_name[c(41:50)], country = NULL, ssl.verifypeer = T )
gendered_names_6 <- genderizeAPI(names$first_name[c(51:60)], country = NULL, ssl.verifypeer = T )
gendered_names_7 <- genderizeAPI(names$first_name[c(61:70)], country = NULL, ssl.verifypeer = T )
gendered_names_8 <- genderizeAPI(names$first_name[c(71:80)], country = NULL, ssl.verifypeer = T )
gendered_names_9 <- genderizeAPI(names$first_name[c(81:90)], country = NULL, ssl.verifypeer = T )
gendered_names_10 <- genderizeAPI(names$first_name[c(91:100)], country = NULL, ssl.verifypeer = T )

#let's look at the names and see if we can find some edge-cases: aka names that might not be 
#decidedly male or female, so we can check them, and change them by hand if necessary
gendered_names_1
gendered_names_2 #edgecase: antonie
gendered_names_3 #edgecase: deni
gendered_names_4
gendered_names_5
gendered_names_6
gendered_names_7
gendered_names_8
gendered_names_9
gendered_names_10
```

The thing with GenderizeR, however, is that one is -- as previously noted -- limited to a 100 names a day, at least, in the free version. Paying for the possiblity to assess the gender of more names was something that I briefly considered, but it eventually looked like a waste of money -- because we can also write our own function.\
\
Well ... other people can write a function we can use. @nielsvullings has created the following function, which can be used to assign a gender score to each name, based on Meertens name-repository.
